Last glacial cycle glacier erosion potential in the Alps

### 2021-02-11 articles/erosion/alps-erosion.tex
Add corrections from Ian, two things to discuss.

    The ice-sheet model set-up was presented in an earlier publication
    \citep{Seguinot.etal.2018} and is briefly summarized here. The simulations
    were conducted using the Parallel Ice Sheet Model (PISM, development
    version~e9d2d1f), an open-source, finite-difference, shallow-ice and
    shallow-shelf model
    \citep{PISM-authors.2017}. Our setup includes temperature and water-content
    dependent creep, pseudo-plastic basal sliding accounting for till
    dilatation under high water pressure, bedrock deformation under the ice
    load, and a~positive degree-day (PDD) surface mass balance model. The model
    is initialized with assumed present-day ice thickness and equilibrium
    ice and bedrock temperature at 120\,ka (thousand years before the present),
    and ran to the present. The ice-sheet model physical parameters are listed
    in the earlier open-access publication \citep{Seguinot.etal.2018}, and the
    complete list of PISM
    parameters is stored in long-term archived model output metadata for the
    present study \citep{Seguinot.2021}, and for the earlier publication
    \citep{Seguinot.2020, Seguinot.2020a}.

@@

    It may seem surprising that the erosion law calibrated on Patagonian and
    Antarctic Peninsula tidewater glaciers \citep{Koppes.etal.2015} yields a
    more realistic glacial cycle erosion potential in the Alps than those based
    on terrestrial
    glaciers \citep{Humphrey.Raymond.1994, Herman.etal.2015, Cook.etal.2020},
    themselves resulting in kilometre-scale integrated erosion even
    higher than the Alpine relief itself. However, during the Last Glacial
    Maximum and much of the last glacial cycle, Alpine paleoglaciers were
    closer in size, slope, and climatic context to the present-day glaciers of
    Patagonia and the Antarctic Peninsula \citep{Koppes.etal.2015} than to
    Franz-Joseph Glacier \citep{Herman.etal.2015} and many of the glaciers
    included in the global compilation by \citet{Cook.etal.2020}. Indeed,
    atmospheric
    temperatures have been found to be a strong driver of erosion in studies
    involving several glaciers \citep{Koppes.etal.2015, Cook.etal.2020}.

@@

    \begin{figure}[ht]
      \centerline{\includegraphics{alpero_basaldrag}}
      \caption{%
        \textbf{(a)} Modelled basal and surface topographies interpolated along
        the Rhine Glacier (using the same transect and ages as in
        Fig.~\ref{fig:basaldrag}).
        \textbf{(b)} Interpolated yield stress in the pseudo-plastic sliding law
        (dotted lines) and magnitude of the basal shear stress (basal drag,
        solid lines), both normalized against the ice overburden pressure
        \citep[cf.][for model physics]{Seguinot.2014, Seguinot.etal.2016}.}
      \label{fig:basaldrag}
    \end{figure}

>>>>

    \todo{Ian: Given the importance of the sliding processes and dependence on
          basal shear stress. Is it worth the thought that "experience with the
          model shows that in the conditions applied here, till is mostly
          saturated and the model behaves similar to Weertman"?}

    The ice-sheet model set-up was presented in an earlier publication
    \citep{Seguinot.etal.2018} and is briefly summarized here. The simulations
    were conducted using the Parallel Ice Sheet Model (PISM, development
    version~e9d2d1f), an open-source, finite-difference, shallow-ice and
    shallow-shelf model \citep{PISM-authors.2017}. Our setup includes
    temperature and water-content dependent creep, pseudo-plastic basal
    sliding, bedrock deformation under the ice
    load, and a~positive degree-day (PDD) surface mass balance model. The model
    is initialized with assumed present-day ice thickness and equilibrium
    ice and bedrock temperature at 120\,ka (thousand years before the present),
    and ran to the present. The ice-sheet model physical parameters are listed
    in the earlier open-access publication \citep{Seguinot.etal.2018}, and the
    complete list of PISM
    parameters is stored in long-term archived model output metadata for the
    present study \citep{Seguinot.2021}, and for the earlier publication
    \citep{Seguinot.2020, Seguinot.2020a}.

@@

    It may seem surprising that the erosion law calibrated on Patagonian and
    Antarctic Peninsula tidewater glaciers \citep{Koppes.etal.2015} yields a
    more realistic glacial cycle erosion potential in the Alps than those based
    on terrestrial
    glaciers \citep{Humphrey.Raymond.1994, Herman.etal.2015, Cook.etal.2020},
    themselves resulting in kilometre-scale integrated erosion even
    higher than the Alpine relief itself. However, during the Last Glacial
    Maximum and much of the last glacial cycle, Alpine paleoglaciers were
    closer in size, slope, and climatic context to the present-day glaciers of
    Patagonia and the Antarctic Peninsula \citep{Koppes.etal.2015} than to
    Franz-Joseph Glacier \citep{Herman.etal.2015} and many of the glaciers
    included in the global compilation by \citet{Cook.etal.2020}.

@@

    \begin{figure}[ht]
      \centerline{\includegraphics{alpero_basaldrag}}
      \caption{%
        \textbf{(a)} Modelled basal and surface topographies interpolated along
        the Rhine Glacier (using the same transect and ages as in
        Fig.~\ref{fig:basaldrag}).
        \textbf{(b)} Interpolated yield stress in the pseudo-plastic sliding law
        (dotted lines) and magnitude of the basal shear stress (basal drag,
        solid lines), both normalized against the ice overburden pressure
        \citep[cf.][for model physics]{Seguinot.2014, Seguinot.etal.2016}.
        Basal drag values greater than yield stress result in glacier sliding.
        \todo{Ian: make sure the legend matches the graph.}}
      \label{fig:basaldrag}
    \end{figure}
